---
title: "Day 4 Homework"
author: "Carey Kopeikin and Steph Neul"
date: "8/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

###Section 1	

1. Create a vector called ```letters.in.name``` that has the number of letters in your first name, your middle name and last name. This would be my answer:
```{r}
#Carey Jason Kopeikin
letters.in.name <- c(5, 5, 8)
```

2. Create a logical vector that contains TRUE for each of your names if your name is longer than 5 letters and FALSE if it is shorter. Do this in the two ways explained below. Note that the output for my names would be:
[1] FALSE FALSE  TRUE

  a. Method 1: create the vector ```is.longer.than.5.method.1``` and for each of your names manually type in TRUE if it is longer than 5 and FALSE if not (there should be a TRUE or FALSE for each of the names). Then call the vector.

  b. Method 2: create the vector ```is.longer.than.5.method.2``` and use logical tests to evaluate if each of your names are longer than 5. Then call the vector.
  
  c. Check to make sure your outputs from parts a and b are the same.
  

3. Create the vector ```some.numbers``` and assign it these numbers 4, 8, 1, -3, 22. Then:

  a.	sort the numbers from biggest to smallest.

  b.	sort the numbers from smallest to biggest.

  c.	find the number of elements in the vector using R.

  d.	display only the second element

  e.	display the 3rd and last elements

  f.	use logical statements to show only the numbers less than 5.

###Section 2 

4. Read in the file `mariokart.csv` and save it as a data frame. Information about `mariokart.csv` can be found here: <https://www.openintro.org/data/index.php?data=mariokart>
You will need to go to that website to read about the data for some of these questions!


5. Display the first 6 observations.


6. Display the first 15 observations:


7. Display the structure of the data frame:


8. Display the number of observations (i.e. how many rows!)


9. Display the number of variables


10. What information does the variable wheels provide? You will need to read about the data on the website provided in order to answer this question. 

###Section 3

11. Ebay takes 10.2% of the total price of the games as a sales fee: <https://www.ebay.com/help/selling/fees-credits-invoices/selling-fees?id=4364>. Create a new column in the data frame that shows adjusted revenue for the seller after Ebay takes out its fee. Give your new column a name that makes sense.

12. Ebay also charges a \$0.35 listing price per item. Create a new column that shows total revenue for the seller after Ebay takes out both its sales fee and listing price. Give your column a new name that makes sense.

